Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-24 paper.qmd
Updated Figure 9 (b)

# Fit high-dimensional model
scurve_model <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = effective_bin1_scurve,
  r2 = r2,
  q = 0.16,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

@@

```{r}
#| label: errors-scurve
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
#buffer_vec <- seq(0.05, 0.1, by = 0.01)

@@

a1_m_scurve <- ggplot(error_scurve,
                     aes(x = a1,
                         y = b_non_empty)) +
  geom_vline(xintercept = 0.12,
         linetype=2,
         color = "#bdbdbd",
         linewidth=1) +
  geom_hline(yintercept = 52,
           linetype=2,
           color = "#bdbdbd",
           linewidth=1) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) +
  # geom_vline(xintercept = 11, linetype="solid",
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 9, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  # geom_vline(xintercept = 13, linetype=2,
  #            color = "#bdbdbd", linewidth=1) +
  scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("non-empty bins (", m, ")"))) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

mse_scurve_b + a1_m_scurve +
  mse_scurve_lwd + 
  plot_layout(ncol=3)
```

>>>>

# Fit high-dimensional model
scurve_model <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = effective_bin1_scurve,
  r2 = r2,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

@@

```{r}
#| label: errors-scurve
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 5:22 #sqrt(NROW(umap_scurve_scaled)/r2)
#buffer_vec <- seq(0.05, 0.1, by = 0.01)

@@

error_scurve <- error_scurve |>
  mutate(prop_bins = b_non_empty/b)

a1_m_scurve <- ggplot(error_scurve,
                     aes(x = a1,
                         y = prop_bins)) +
  geom_vline(xintercept = 0.12, linetype="solid",
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 0.10, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 0.18, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) +
  # scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

mse_scurve_b + a1_m_scurve +
  mse_scurve_lwd + 
  plot_layout(ncol=2)
```
